\mainmatter

# Introduction

**Before** you start to use jamovi or SPSS, you need to know *why* you need the software.

* If your RQ concerns the **relationship between two qualitative variables**:
  - Analysis of **odds ratios** using, for example, a $\chi^2$-test; or
  - Analysis of **two proportions** using, for example, a $z$-test.
  - Displays using a side-by-side or stacked barchart.

* If your RQ concerns the **difference between means in two groups**:
  - Analysis of two means using, for example, a two-sample $t$-test.
  - Displays using a error bar chart (and perhaps a boxplot).

The data files used in this document are available in Sect. \@ref(DataFiles).

You can download the data, and follow the instructions in this document.





## The data: two qualitative variables {#B12Data}

Consider data [@data:Gammon2012:B12] [in Exercise\ 31.16 of the textbook](https://peterkdunn.github.io/SRM-Textbook/AnalysisOddsRatio.html) (Table\ \@ref(tab:jamoviB12Data)).



```{block2, type="rmddatafile"}
The data file (`B12-Long.csv`) can be downloaded in Sect. \@ref(DataFiles).
You can download the data, and follow the instructions.
```


**This data set is available from the textbook as `B12-Long`** in 
[Appendix\ A](https://peterkdunn.github.io/SRM-Textbook/AppendixDataSets.html).
You can download the data and follow along if you wish.

There are *two qualitative variables* being measured on each individual:

1.  Whether the person is vegetarian (or not);
1.  Whether the person is B12 deficient (or not).

Hence, jamovi will have *two columns*: one for each variable.
There are $124$ units of analysis, so there are *$124$ rows* of data.


```{r jamoviB12Data,  echo=FALSE}
B12Data <- array( dim = c(3, 3) )

rownames(B12Data) <- c("Vegetarians",
                       "Non-vegetarians",
                       "Total")
colnames(B12Data) <- c("B12 deficient",
                       "Not B12 deficient",
                       "Total")

B12Data[1, ] <- c(8, 26, 34)
B12Data[2, ] <- c(8, 82, 90)
B12Data[3, ] <- c(16, 108, 124)

if( knitr::is_latex_output() ) {
  kable(B12Data,
        format = "latex",
        booktabs = TRUE,
        longtable = FALSE,
        align = c("r", "r", "r"),
        caption = "The number of vegetarian and non-vegetarian women who are (and are not) B12 deficient.") %>%
    row_spec(0, 
             bold = TRUE) %>%
    row_spec(3, 
             bold = TRUE) %>%
    column_spec(4, 
                bold = TRUE) %>%
    row_spec(2, hline_after = TRUE) %>%
    kable_styling(font_size = 10)
  
}
if( knitr::is_html_output() ) {
  kable(B12Data,
        format = "html",
        booktabs = TRUE,
        longtable = FALSE,
        align = c("r","r", "r"),
        caption = "The number of vegetarian and non-vegetarian women who are (and are not) B12 deficient.") %>%
    row_spec(0, 
             bold = TRUE) %>%
    row_spec(3, 
             bold = TRUE) %>%
    column_spec(4, 
                bold = TRUE)
}
```



## The data: one qualitative, one quantitative variable {#jamoviLeanAngle}

Consider the face-plant data [@data:Wojcik:ForwardFall] [used in Exercise\ 30.11 of the textbook](https://bookdown.org/pkaldunn/SRM-Textbook/AnalysisTwoMeans.html) (Table\ \@ref(tab:jamoviFacePlant)).



```{block2, type="rmddatafile"}
The data file (`ForwardFall.csv`) can be downloaded in Sect. \@ref(DataFiles).

You can download the data, and follow the instructions.
```


```{r jamoviFacePlant, echo=FALSE}
FacePlant <- cbind( 
  c(29, 32, 34, 31, 33),
  c(34, 27, 32, 28, 27),
  c(18, 15, 23, 13, 12) 
)

if( knitr::is_latex_output() ) {
  kable( FacePlant,
         format = "latex",
         longtable = FALSE,
         booktabs = TRUE,
         align = c("c", "c", "c"),
         #col.names=rep("", 3),
         caption = "Lean- angles for older and younger women.") %>%
    add_header_above(header = c("Younger women " = 2, 
                                "Older women " = 1), 
                     bold = TRUE, 
                     align = "c") %>%
    kable_styling(font_size = 10)
  
}
if( knitr::is_html_output() ) {
  kable( FacePlant,
         format = "html",
         longtable =FALSE,
         booktabs = TRUE,
         align = c("c", "c", "c"),
         col.names = rep("", 3),
         caption = "Lean- angles for older and younger women.") %>%
    kable_styling(full_width = FALSE) %>%
    add_header_above(header = c("Younger women " = 2, 
                                "Older women " = 1), 
                     bold = TRUE, 
                     align = "c")
}
```


